### 2011-11-23 paper/drinking.tex

\documentclass{bioinfo}
\usepackage{setspace}

\copyrightyear{2005}
\pubyear{2005}
\usepackage{url}
\usepackage[normalem]{ulem}
\usepackage{multirow}
\usepackage{pifont}
% Load packages
\usepackage{cite} % Make references as [1-4], not [1,2,3,4]
\usepackage{url}  % Formatting web addresses  
\usepackage{ifthen}  % Conditional 
\usepackage[utf8]{inputenc} %unicode support
\urlstyle{rm}
\usepackage{natbib}
\renewcommand{\cite}{\citep}
\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{pifont}
%\newcommand{\tick}{\ding{52}}
\newcommand{\tick}{\checkmark}
\newcommand{\tickNo}{\hspace{1pt}\ding{55}}

\begin{document}
\firstpage{1}

\title{Process-based phenotypes: an application to the Cellular
  Phenotype Ontology}

\author[Hoehndorf et al.]{Robert Hoehndorf$^1$ and Gabriella
  Rustici$^2$ and Midori A. Harris$^3$ and Georgios V. Gkoutos$^{1}$}

\address{$^{1}$Department of Genetics, University of Cambridge,
  Downing Street, Cambridge, Cambridge CB2 3EH, UK\\
  $^{2}$EBI, Wellcome Trust Genome Campus, Hinxton, Cambridge,
  Cambridge CB10 1SD, UK\\
  $^{3}$Department of Biochemistry; University of Cambridge, 80 Tennis
  Court Road, Cambridge CB2 1GA, UK}

\maketitle

\begin{abstract}
  Phenotypes important for...

  Include behavioral phenotypes, based on behavior ontology

  We identify a problem in representing phenotypes such as drinking
  behavior that...

  provide solution...
\end{abstract}

{\bf Keywords:} cell phenotype, process, ontology, Semantic Web

\section{Introduction}

problem of phenotypes...

why reasoning, why formal, success,...

solved: structures, qualities

open: process phenotypes very complex...  
physiology...
measurements...

\subsection{Problem statement 1}
Heart rate increased: frequency of CMC increased fluid flow increased
volume per contraction increased ...

what is the relation to the heart?

\subsection{Problem statement 2}
increased apoptosis

\subsection{Observations}
observation 1: there is a fundamental distinction between attribution
of qualities for a fixed period (e.g., 24h) and single {\em act}s of
drinking.

observation 2: each of the different possible ``increased'' drinking
behavior may be the result of different underlying causes and due to
this, distinguishing between them may be important.

observation 3: depending on the experiment performed, only some of the
actual qualities can be measured.

observation 4: some qualities can be inferred given others. For
example,...

\section{Materials and Methods}
\subsection{Formal ontology}
The ontology as an approach to semantic standardisation was proposed
more than a decade ago and since then has become the dominant
methodology used to semantically categorise phenodeviance.  The
biomedical research community has invested considerable amount of
effort and resources in the development and establishment of
ontologies that are becoming increasingly successful as information
management and integration tools in many disparate scientific fields
allowing interoperability and semantic information processing between
diverse biomedical resources and domains.

In computer science, an ontology is a specification of a
conceptualization of a domain of knowledge \cite{Gruber1995,
  Guarino1998}.  Ontologies commonly distinguish between {\em classes}
and {\em individuals} within a domain of knowledge. A class is an
entity that can have {\em instances}, while individuals are entities
that cannot be instantiated. Examples of individuals include the
Eiffel tower or the 2009 Ironman Triathlon in Hawaii, while examples
of classes include {\em Tower} or {\em Triathlon}. The Eiffel tower
can be an instance of the class {\em Tower}, and the 2009 Ironman
Triathlon an instance of {\em Triathlon}.  The meaning of classes is
specified by stating what must be true of their instances.

In addition to classes and individuals, ontologies often include {\em
  relations}. Relations are the ``the glue that holds things together,
the primary constituents of the facts that go to make up reality''
\cite{}.

In {\em formal} ontologies, the specification of classes and relations
follows the axiomatic-deductive method. Given a set of terms that are
used within a domain and whose meaning we wish to specify, we begin by
providing {\em explicit definitions} for some terms, potentially
introducing new terms. An explicit definition of a term $t$ is a
statement that can replace every occurrence of $t$ in any sentence.

Eventually, a set of {\em primitive terms} remains that are not
further defined. Using only the primitive terms, we can construct
complex sentences. Based on the intended meaning of the primitive
terms, we consider some of these sentences true and some of them false
in our domain. We select some of the true sentences as {\em axioms}
which provide the core of our ontology. Ideally, the axioms are chosen
so that all true sentences in the domain we intend to represent follow
by means of logical deduction from the axioms. More commonly, however,
only {\em some aspects} of the intended meaning are formally
represented while other aspects are omitted either due to limitations
in language expressivity or due to their irrelevance to the problem
for which an ontology is developed.

Based on the axioms and definitions, we can use deduction to infer
statements that logically follow from the axioms.  The process of
automatically deducing sentences from axioms is called {\em automated
  reasoning}. Automated reasoning allows users of an ontology to carry
our key activities: verifying the ontology's consistency, inferring
hidden knowledge and thereby performing powerful queries.  An ontology
is inconsistent if {\em every} statement can be inferred from the
ontology. In particular, an ontology is inconsistent if a statement
$\phi$ and its negation $\neg \phi$ can be inferred from the
ontology's axioms. Classes in an ontology may be unsatisfiable: a
class $C$ is unsatisfiable, if it is not possible that the class has
instances. Unsatisfiable classes are commonly the result of a
contradictory class definition. That a class is unsatisfiable is often
a special kind of unintended consequence that can be drawn from an
ontology.

Automated reasoning in the Web Ontology Language (OWL) can be employed
to automatically compute the generalization hierarchy underlying an
ontology as well as for verification of data consistency and complex
queries \cite{Hoehndorf2011incon, Hoehndorf2011models}.

Layered approach: although OWL has high expressivity, some things
cannot be expressed in OWL; some things should not be expressed in OWL
for specific applications due to tractability reasons

Most applications rely on graph structure of ontology...

\subsection{Processes, their participants and qualities}
Processes are entities that are temporally extended: processes have
temporal parts and {\em unfold} through time. Examples of classes that
may have processes as instances include {\em Drinking}, {\em
  Triathlon} or {\em Apoptosis}.

Processes commonly have participants\footnote{It may be argued that
  processes necessarily have participants or that processes may have
  no participants. However, these discussions do not affect our aims
  here and take no position in this regard.}. Participants of
processes are entities that are present at time points (i.e., they
have no temporal parts) and may undergo changes during the process.
We can further distinguish different {\em roles} of participation in a
process. For example, a runner may participate in a {\em Triathlon}
process as a {\em Runner (role)}, while a particular {\em goal line}
can participate in the same process in a different role.

Amongst the participants of some processes, we will distinguish
between {\em inputs} and {\em outputs} of processes. Chemical
reactions, for example, are processes with definite inputs (the {\em
  reactants}) and outputs ({\em the products}).

\subsection{Gene Ontology}
quick intro...

mention regulation relations...

regulation of X processes

\subsection{PATO and the EQ model}

\section{Results}
\subsection{Attributes of processes}
An attribute that can be directly linked to a process is its {\em
  duration}. A duration is an attribute that characterizes the
temporal extent of a process and is similar to {\em Length}, {\em
  Area} and {\em Volume} for one-, two- and three-dimensional
entities. Furthermore, processes may be attributed a {\em temporal
  location} with respect to a particular coordinate system.

However, the majority of attributes that characterize processes are
not based on genuine process attributes alone, but rather relate
attributes of process participants with the duration of a process. In
particular, a {\em rate} typically refers to an attribute of some
entity {\em with respect to an attribute of another entity}, and in
the context of processes, rates often refer to attributes of a process
participant with respect to the duration of the process. For example,
a {\em mass flow rate} refers to the {\em Mass} of a process
participant with respect to the duration of the process, i.e., how
much matter is moved (from one point to another) through the process.
As a more complex example, a {\em rate of change of position} refers
to the {\em distance} that an object is moved with respect to the
duration of the process.

Not all rates of a process depend on attributes of the process
participants. In particular, a {\em frequency of occurrence} or {\em
  event rate} refers to the number of occurrences of a type of process
during the reference process. For example, a {\em rate of heart
  beating} refers to the number of {\em Heart beating} processes that
occur within a reference process (e.g., a process containing the heart
with a duration of one minute). Further attributes that depend on
types of processes with regard to a reference process are {\em
  distribution patterns}, i.e., how the occurrences of processes of a
particular type are distributed within a reference process. For
example, the heart may beat {\em rhythmically} or {\em arrhythmically}
within a period of time.

Related to distribution patterns are {\em changing qualities} of
processes. For example, the rate of heart beating may change ({\em
  increase} or {\em decrease}) throughout the course of the reference
process. A simple analysis of {\em increasing} ({\em decreasing})
rates would be that the rate of a heart beating within the first half
of a process if {\em lower} ({\em higher}) than in the second half of
the process. Mathematically, this process sub-division can be iterated
until we end up with infinitesimally small time intervals.

While some processes can be subdivided indefinitely without changing
the type of process, others cannot. A class $C$ of {\em continuous
  processes} is a class that has processes $p$ as instance such that
all temporal parts of $p$ are also instances of $C$. Examples include
{\em continuous movements} or {\em regulation} processes. On the other
hand, {\em discrete processes} can be subdivided into stages of
activity and stages of inactivity: a class $C$ of {\em discrete
  processes} is a class with processes $p$ as instance and there is a
part of $p$ that is not an instance of $C$. An example is {\em heart
  beating} which has periods of activity (a single heart beat) and
inactivity.

% In line with the GO, we call processes that cause

% [Regulation of X vs. X: regulation modulates the number/amount of
% occurrence...]

We may say that an object has a {\em frequency} or {\em rate} as if
these were attributes of objects at time points. For example, that my
heart {\em now} beats with a frequency of 80 beats per minute, or that
a car is moving with a speed of 280 kilometres per minute {\em at a
  particular point in time} (e.g., as observed with a speed camera)
may be considered attributes of the objects, not attributes of
processes. In our analysis, rates as attributes of objects can be
explicitly defined using rates of processes. For example, the heart
beating frequency of a particular heart $h$ at a time point $t$ is the
frequency of heart beating processes in which $h$ participates at an
infinitesimally small process around $t$ and within a reference
process. The reference process is necessary in order to obtain a value
for a frequency or rate even if the process in question is not
occurring. For example, a rate of cardiac muscle contraction can be
determined even when no cardiac muscle contraction is occurring in
virtue of a reference process with a rate of cardiac muscle
contraction.

The reference process for an object at a time point $t$ is uniquely
determined for continuous processes (the infinitesimally small process
around $t$), and ambiguous for discrete processes.

A similar construction as for continuous processes can be made for
discrete processes. A rate of blood flow is immediately an attribute
of a single heart beat process. However, during a period of time, the
cumulative rate of blood flow can be observed as the sum of the blood
flow rates of individual heart beating processes.

\begin{figure}
  \centering
  \includegraphics[width=.5\textwidth]{processpatterns.pdf}
  \caption{Overview over ...\label{fig:patterns}}
\end{figure}

Complex relations between Based on these definition patterns,
interdependencies between process attributes can be observed.

\subsection{Cellular Phenotype Ontology}

A domain that would greatly benefit from the development of a CPO is
Systems Microscopy, which aims to understand complex and dynamic
cellular systems by combining automated fluorescence microscopy, cell
microarray platforms, quantitative image analysis and data mining
(Lock \& Stromblad, Exp. Cell Res. 2010).

If we consider some of the studies, which have been published in this
field in the last few years (Neumann et al, Nature 464: 722-727;
Schmitz et al, Nature Cell Biology 12: 886-893; Fuchs et al, Mol Sys
Biology 6: 370), the need for a CPO becomes evident.

In all three studies here considered live-cell imaging assays and RNAi
knockdown were used to generate phenotypic profiles that quantify the
cellular response to a given siRNA thus allowing identification of
hundreds of genes involved in diverse biological functions including
cell division, migration and survival.

In each study, several phenotypes were detected and described by the
authors without the use of ontologies (see table), making the
integration between datasets extremely difficult. For example, it is
evident that cell division phenotypes were observed in the three
datasets (e.g. 'Mitotic delay/arrest' and 'Prolonged mitotic exit' or
'Methaphase delay' and 'Methaphase cells') but the overlap between
such phenotypes is unclear.

Data integration is also complicated by the lack of standardization at
the level of data production and processing; all these issues are
currently being address by the different groups involved in the
Systems Microscopy Network of Excellence
(http://www.systemsmicroscopy.eu/) and the first step towards data
integration will be achieved by developing a CPO.

This ontology will be used to integrate phenotypes' definitions across
existing datasets and will then become an integrated part of the data
processing pipeline and used to annotate the data as it gets generated
(Conrad et al, Nature Methods 8: 246-249).

\begin{figure*}[h]
  \centering
  \includegraphics[width=\textwidth,
  height=.33\textheight]{overview.pdf}
  \caption{Overview over ...\label{fig:overview}}
\end{figure*}

CPO contains both structural abnormalities and physiological
abnormalities.

Phenotypes in the CPO are either based on structural abnormalities or
abnormal physiology involving cells or cell components. Structural
abnormalities in the CPO are based on GO's Cellular Component (GO-CC)
hierarchy. GO-CC contains 2,918 classes for cell parts (including {\em
  Cell}) and extracellular components of cells. For each cellular
component class $C$ in the GO-CC, we create a new class labelled {\em
  $C$ phenotype} in the CPO. For example, for the class {\em
  Mitochondrion} ({\tt GO:0005739}) in the GO-CC, we create the class
{\em Mitochondrion phenotype}.

Amongst the structural phenotype classes, we first distinguish between
{\em normal} and {\em abnormal} phenotypes. An {\em Abnormal phenotype
  of $C$} is a phenotype of an organism that does not have a normal
$C$ as part, while a {\em Normal phenotype of $C$} represents the
state in which an organism has a {\em normal $C$} as part.

Amongst the abnormal phenotypes that we include for all cell
components listed in GO-CC, we distinguish {\em abnormal morphology}
and {\em abnormal physiology} phenotypes. An {\em Abnormal morphology
  of $C$} is either the (abnormal) absence of required parts of $C$,
the (abnormal) presence of additional parts, or abnormal qualities of
$C$ \cite{Hoehndorf2010phene}. For example, an absence of caveolae
({\tt MP:...}) would be a subclass of {\em Abnormal morphology of
  plasma membrane} in virtue of caveolae necessarily being part of the
{\em Plasma membrane} ({\tt GO:0005886}).

Abnormal physiology of a cell component refers to abnormal {\em
  functionality} of a cell component. We assume that a functionality
of a cell component is (the potential for) a process in which the cell
component is (causally) involved. We use the definitions of GO classes
that were created based on lexical decompositions of GO class labels
\cite{Mungall2010go, Bada2007a, Ogren2004} to identify the processes
in which cell components are involved. For example, the definition of
the GO class {\em Mitochondrial fission} ({\tt GO:0000266}) is
explicitly defined as an {\em Organelle fission} ({\tt GO:0048285})
that {\bf results-in-the-division-of} a {\em Mitochondrion} ({\tt
  GO:0005739}). Based on this definition, we make the assumption that
{\em Mitochondrial fission} is one of the functions of a {\em
  Mitochondrion} and that an {\em Abnormality of mitochondrial
  fission} is a subclass of an {\em Abnormality of mitochondrion
  physiology}.

Amongst abnormal physiology, we distinguish between abnormalities in a
{\em single occurrence} of a cell component's functioning and an
abnormal {\em pattern of multiple occurrences} of a cell component's
functioning. For example, abnormalities in cell division resulting in
{\em Aneuploidy} refer to abnormalities of functioning processes,
while an {\em increased rate of cell division} refers to an
abnormality in the pattern of occurrence of multiple cell division
processes. In the CPO, abnormalities in the pattern of occurrence of
$X$ are represented as abnormalities in the {\em regulation of $X$}.

Single occurrences of processes can be abnormal in multiple ways,
depending on the type of process.
%
First, common to all processes is the quality of {\em duration} and
consequently, each process can have an {\em abnormal} (increased or
decreased) duration. For example, a part of an organism may
participate in an {\em Inflammatory response} ({\tt GO:0006954}) that
lasts longer than normal, i.e., the organism has an {\em Increased
  duration of inflammatory response} phenotype. We define such a
phenotype as a phenotype of an organism which has a part that
participates in {\em Inflammatory response}, and this {\em
  Inflammatory response} process has an {\em Increased duration} ({\tt
  PATO:0000498}).

The second common type of abnormality are abnormalities based on
process participants in relation to the duration of the process. These
include all kinds of {\em rates} such as {\em mass flow rate}, {\em
  energy flow rate} and {\em velocity} (the rate of change of
position). In each of these cases, an object participates in a process
and a quality (or change of quality) of that object throughout the
duration of the process is considered to form a new quality. If the
process has participants that are distinguished into {\em inputs} and
{\em outputs}, then a recurring pattern is that the amount of inputs
or outputs that participate in the process can be {\em increased} or
{\em decreased}. For example, an {\em Increased rate of cytoplasmic
  streaming} can be defined as an increased amount of inputs or an
increased amount of outputs of a {\em cytoplasmic streaming} process.

Finally, some objects may be divided into stages during which
particular states of affairs obtain, and a process may be abnormal in
that these states of affairs do not obtain at a particular
stage. Notably, at the beginning and the end of a process, pre- and
post-conditions may obtain that are abnormally changed in a
process. For example, {\em Aneuploidy} -- an abnormality during cell
division at which the chromosomes do not separate properly between the
two cells -- may be considered the result of such an abnormality.

We implement the first two types of abnormality in the CPO. First, as
a subclass of each {\em Abnormality of P} class, we create {\em
  Abnormal duration of P}, which in turn has {\em Increased duration
  of P} and {\em Decreased duration of P} as subclasses. Second, if we
are able to identify {\em inputs} $I(P)$ or {\em outputs} $O(P)$ of
the process $P$ in the formal definitions of the GO, we automatically
generate {\em Abnormality of $I(P)$ in $P$} as well as {\em
  Abnormality of $O(P)$ in $P$}.  The left side of Figure
\ref{fig:overview} illustrates the schema of classes we generate for
single process abnormalities.

The second type of abnormality in the CPO relate to abnormalities of
{\em multiple} occurrences of some process $X$. According to the GO,
{\em regulation} processes are processes that maintain or modify the
occurrence of processes of a particular type. Following this
convention, we call an abnormality of multiple occurrences of $X$ {\em
  abnormality of the regulation of $X$}.

A first kind of abnormality of regulatory processes are {\em abnormal
  temporal distribution patterns} of a process. In these
abnormalities, the {\em way} in which processes of a particular kind
are temporally distributed is abnormal.  The most common abnormal
distribution pattern is an increased or decreased frequency, and we
use PATO's {\em frequency} class to define {\em Abnormal frequency of
  occurrence of $X$}.
% \begin{verbatim}
% has-phenotype some (has-part-some (participates-in some 
%   (regulates some X and has-quality some (frequency and towards some X))))
% \end{verbatim}
For example, an {\em Abnormal frequency of occurrence of apoptosis} is
defined as an abnormality of {\em Regulation of apoptosis} ({\tt
  GO:0042981}) with respect to the {\em frequency} ({\tt
  PATO:0000044}) of {\em Apoptosis} ({\tt GO:0006915}) occurrences.

There are further types of deviation from a distribution pattern. For
example, a kind of process that is normally {\em rhythmic} can be
abnormal in that it is {\em arrhythmic}. A typical example of this
kind of process is {\em Heart beating} ({\tt GO:0060047}), in which
{\em Cardiac muscle contraction} ({\tt GO:0060048}) processes occur in
a rhythmic pattern. In {\em Cardiac dysrhythmia}, however, {\em
  Cardiac muscle contraction} processes occur arrhythmically, and we
consider this to be an abnormality of the regulation of {\em Cardiac
  muscle contraction}. While these abnormalities are often highly
informative in clinical diagnostics and biological investigations, we
usually lack the necessary information that is required to
automatically determine meaningful types of abnormal distribution
patterns.

A second kind of regulatory abnormalities is related to the {\em
  onset} of a process. With respect to a reference process, a
particular kind of process may be {\em delayed} ({\tt PATO:0000502})
or {\em premature} ({\tt PATO:0000694}). For example, {\em Delayed
  apoptosis} refers to an abnormality of the {\em Regulation of
  apoptosis} in which apoptosis is induced later than normal.  We use
the PATO quality {\em onset} ({\tt PATO:...}) and its children {\em
  delayed} and {\em premature} to define these types of regulatory
abnormality. Similarly, we use PATO's {\em offset} ({\tt PATO:...})
  quality and its children to characterize regulatory abnormalities in
  which a process ends prematurely or too late.

\begin{table*}
  \centering
  \begin{tabular}{l|l|l|l}
    & Increased cytoplasmic flow rate & Normal cytoplasmic flow rate & Decreased
    cytoplasmic flow rate \\
    \hline
    Increased frequency &increased total flow rate &increased total
    flow rate &?\\
    Normal frequency &increased total flow rate &normal total flow
    rate &decreased total flow rate\\
    Decreased frequency &?&decreased total flow rate &decreased total
    flow rate\\
    \hline
  \end{tabular}
  \caption{\label{tbl:flow}Total cytoplasmic flow rate...}
\end{table*}

Finally, a third kind of regulatory abnormality refers to abnormal
rates with respect to a participant of the process that is being
regulated. For example, a cytoplasmic flow rate can be increased or
decreased not within a single {\em cytoplasmic streaming} process but
rather the total cytoplasmic flow rate, as a summation over all
cytoplasmic streaming processes that occur within an organism (or a
particular anatomical location), is increased or decreased. While a
flow rate of a single {\em cytoplasmic streaming} process is a quality
of that process, an increased {\em total} cytoplasmic flow rate is a
quality of the regulation of {\em cytoplasmic streaming}. In
particular, it is possible for an organism to have a normal --- or
even a decreased --- cytoplasmic flow rate in each individual
cytoplasmic streaming process while at the same time having an
increased total cytoplasmic flow rate due to a large increase in the
frequency of occurrence of cytoplasmic streaming processes. Similarly,
the frequency of occurrence of cytoplasmic streaming may be normal or
decreased while the total cytoplasmic flow rate is increased due to an
increased cytoplasmic flow rate in each individual cytoplasmic
streaming process. Table \ref{tbl:flow} illustrates the dependencies
between rates of individual processes, their frequency of occurrence
and the total rate of these processes. We include total rates as
regulatory abnormalities in the CPO since these are the attributes of
processes that are often measured or observed, while the rates of
individual processes are inferred following a schema such as Table
\ref{tbl:flow}.

\subsection{Cell Phenotype Ontology: implementation}
We were faced with two choices for implementing the CPO: we could
either implement a pre-composed ontology in which all classes and
their definitions are pre-generated according to the patterns we
define, or we could develop an annotation software that enables the
selection of our process phenotype patterns based on the current
structure of the GO.  To maximize the utility and compatibility of the
CPO, and to provide stable identifiers for its concepts, we selected
the first strategy and developed a software to automatically generate
a pre-composed ontology from the GO.

We developed a software that utilizes the OWL API \cite{Horridge2007}
in order to generate an OWL representation of the CPO. The software
requires three input files: a version of the GO on which to based the
generated CPO, a version of PATO that is used to define abnormal
qualities, and a copy of the GO cross-product definitions
\cite{Mungall2010go} that is used to relate cell components to the
processes in which they participate as well as identify the
participants, inputs and outputs of processes.

We automatically generate a unique numerical identifier for each class
in the CPO.  Since the CPO is based on the GO and need to be updated
with subsequent versions of the GO, we must ensure to keep identifiers
stable in subsequent versions of CPO. Therefore, we use the
identifiers for GO classes to generate CPO class identifiers.

In the CPO, identifiers contain two components and are of the form
{\tt CPO:XXGGGGGGG}, where {\tt GGGGGGG} is the seven-digit identifier
of the GO class on which the CPO class is based, and {\tt XX} is a
prefix that identifies the type of phenotype pattern that is applied
to the GO class. For example, based on the class {\em Apoptosis} ({\tt
  GO:0006915}), we generate the CPO classes {\em Abnormality of
  Apoptosis}, {\em Abnormality of single occurrence of apoptosis} and
{\em Abnormality of regulation of apoptosis}.  We use the prefixes
{\tt 12}, {\tt 14} and {\tt 15} for each of the corresponding
phenotype patterns, and consequently generate the class identifiers
{\tt CPO:120006915}, {\tt CPO:140006915} and {\tt CPO:150006915}. As
long as the GO maintains its identifier for the {\em Apoptosis} class,
the identifiers in the CPO will remain stable even when it is
regenerated.

We use the labels of GO classes to automatically generate class labels
for phenotype classes as well as textual definitions for classes in
the CPO. For example, the label of the class for increased number of
occurrences of {\em Apoptosis} is {\em Increased frequency of
  occurrences of Apoptosis}, and its textual definition states: ``An
increased frequency of occurrences of Apoptosis is a phenotype of
entities in which the number of occurrences of Apoptosis within a
given time period is increased in comparison to a reference process.''

\section{Discussion}
\subsection{Implementation as pre-composed ontology}
Pre-composed or software support for pattern?

\subsection{Use of CPO}

\subsection{Future research: acts and intentional states}
activities/actions, intentional (goal-oriented) acts

feeling thirsty (internal state) as driving force behind drinking
behavior

non-measurable, inference only

\subsection{Formalization}
Formal, axiomatic system for process qualities, maybe based on PSL,
FOIS paper 2010

one example:...

\subsection{Conclusion}

\section{Acknowledgements}

% % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section*{Authors contributions}

% % %%%%%%%%%%%%%%%%%%%%%%%%%%%
% The authors declare that they have no competing interests.

% % %%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section*{Acknowledgements}
% \ifthenelse{\boolean{publ}}{\small}{}

% The research work in its first unrevised form was presented at the
% SMBM 2010, Hinxton, Cambridge, U.K.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The Bibliography %%
%%                                                         %%
%% Bmc_article.bst will be used to %%
%% create a .BBL file for submission, which includes %%
%% XML structured for BMC.  %%
%%                                                         %%
%%                                                         %%
%% Note that the displayed Bibliography will not %%
%% necessarily be rendered by Latex exactly as specified %%
%% in the online Instructions for Authors.  %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{plain} % Style BST file
\bibliography{lc}

\end{document}