### 2012-04-02 paper/response-bioinformatics.txt

Dear Dr. Bateman,

Thank you for providing us with the opportunity to revise our
manuscript. We have now prepared a revision in which we have addressed
all reviewers' comments. In response to both reviewers, we have
removed the discussion of the heart beating example from the
introduction and now provide a discussion of the axioms in the
manuscript as well as a detailed presentation of the axioms as
supplementary material. We have included a response to the individual
points raised by the reviewers below.  We would like to thank the
reviewers for their valuable comments which allowed us to improve our
manuscript.

Please do not hesitate to contact us if you require any further
information.

With best wishes,
Robert.

Responses to reviewers
----------------------

Reviewer 1:
"Introduction: I find the heart beating example difficult to follow,
primarily because it lacks simple easy-to-understand definitions - so
either this section is removed or needs to be rewritten (the result
section, however, is much clearer)."

Response: We have removed the section on heart beating from the
introduction, as recommended by the reviewer.

Reviewer 1:
p4 - On the one hand, I agree with "abnormalities in cell division
resulting in Aneuploidy refer to abnormalities of cell division
processes, while an increased rate of cell division refers to an
abnormality in the pattern of occurrence of multiple cell division
processes". However, you then state "an increased rate of cell
division is not an attribute of cell division processes, but rather an
attribute of the regulation of cell division"
two questions: 
1. is the regulation of a process(es) not different from that
collection of processes?
2. is it not the case that a collection of processes has attributes
based on its parts?

Response:
We agree with the reviewer that "regulation" processes do not always
(or commonly) include the processes that are being regulated as
parts, and we have modified our manuscript in this regard, providing
more details on regulation processes in GO and their relation to
collection processes.

Reviewer 1:
This paper needs to discuss the axiom patterns that generates the ontology.

Response:
We agree with the reviewer, and now provide information about the
axioms both in the manuscript (new section 2.3 "Axioms for physiology
phenotypes") and full details as supplementary material.

Reviewer 1:
abstract: suggests that its application is in the unification of
cellular phenotypes across species; but it's not clear where
species-specific differentiation would come into play. it also
suggests (and is further outlined in the discussion) that it could be
ussed for systems microscopy, but no evidence is provided to support
the claim. And the latter, really must be further developed, to
demonstrate that the ontology can be used for some application.

Response: 
The Cell Phenotype Ontology, which we have built by applying our
ontology-patterns for representing physiology phenotypes, is currently
being applied for the annotation of cellular observations in systems
microscopy. For this purpose, CPO provides the necessary terms (in the
sense of a controlled vocabulary) as well as a taxonomic structure
that is used for basic retrieval operations.

We agree with the reviewers that an application of our work in a
bioinformatics analysis would be highly beneficial. For example, once
sufficient data is annotated with the CPO, we can use its taxonomic
structure for enrichment analyses. Similarly, once more cellular
phenotypes are being annotated in various model organism databases, we
can utilize CPO for integrative phenotype analyses across
species. However, at the moment, no comprehensive ontology of
physiology phenotypes is available, partly due to the complexity of
the domain and partly due to limitations in available ontology
frameworks. We see our main contribution in the development of a
theory that allows to capture physiology phenotypes; we applied this
theory to construct (largely automatically) the CPO, which is now
being used to annotate cellular phenotypes in the systems microscopy
domain. Based on our contributions, it is now possible to annotate
and eventually integrate (through the use of a controlled vocabulary
of physiology phenotypes, of which the CPO provides a first example)
data in the domain of physiology. At the same time, we maintain
compatibility with GO (on which a majority of physiology phenotypes
are based) and other PATO-based phenotype ontologies, thereby filling
a gap that has limited data integration in the physiology domain so
far.

Reviewer 2:
A theory of physiology related phenotypes and an implementation based
on GO, GO cross-products and PATO is presented. An outline of a theory
of phenotypes is introduced which participate-in a set of pathological
processes which can be qualified by a small set of PATO qualities. At
least, for a majority of heterogeneous phenotype descriptions this can
be an approach for unification. In future work a sound formal
characterization has and an evaluation of the practical value of the
approach has to be delivered.

Response:
We have now added a detailed formal presentation of our axioms as
supplementary material. An evaluation of the practical utility depends
on data that is being annotated with physiology phenotypes. Currently,
no comprehensive ontology of physiology phenotypes is available. Our
work brings the immediate benefit that the availability of CPO enables
annotation of cellular phenotypes in the systems microscropy
community, and our method can be applied to physiology phenotypes in
other domains, in particular computational modelling.

Reviewer 2:
Minor revision (page 2 line 27): You introduce a Heart beating example
to illustrate the challenges in modeling processes and their
regulation. [...] The Heart example is illustrative but limited with
respect to your ontology. There, both the the function of the single
entity can be impaired and the regulation of multiple occurences of
processes (where the example ends due to the single heartedness of
many animals).

Response: 
In response to this comment and reviewer 1's comment, we have removed
the Heart beating example from the introduction and focus on the
detailed presentation of the processes involved in the Results section
of our manuscript.

Reviewer 2:
Minor revision (page 3 second column line 33) see as above on the
phases of a single Heart action. The heart has no phase of
“inactivity” (that would be death), only a number of phases of which
the filling/ ejection phases are the least granular.

Response: We have changed our manuscript

Reviewer 2:
Minor revision (page 5 Table 1): make it clear that you refer to the
product of row and column quantitites in the cells. I find it
difficult to understand this easy relation from your legend. Do you
really need this table?

Response:
We have removed the table from our manuscript, and discuss the
content of the table now in the main text.

Reviewer 2:
Class names should be consistently written italic and uppercase
throughout the manuscript (e.g. Heart beating and Rate of heart
beating)

Response:
We have made this change throughout our manuscript.